Collaborative Spatial Reuse in Wireless Networks via Selfish Multi-Armed Bandits

### 2018-09-06 OLD SUBMISSION/Article (Elsevier) - LaTeX files/collaborative_spatial_reuse_via_mabs.tex
New + old submission differentiation

\begin{abstract}
Next-generation wireless deployments are characterized by being dense and uncoordinated, which often leads to inefficient use of resources and poor performance. To solve this, we envision the utilization of completely decentralized mechanisms to enable Spatial Reuse (SR). In particular, we concentrate in Reinforcement Learning (RL), and more specifically, in Multi-Armed Bandits (MABs), to allow networks to modify both their transmission power and channel based on their experienced throughput. In this work, we study the exploration-exploitation trade-off by means of the $\varepsilon$-greedy, EXP3, UCB and Thompson sampling action-selection strategies. Our results show that optimal proportional fairness can be achieved, even if no information about neighboring networks is available to the learners and WNs operate selfishly. However, there is high temporal variability in the throughput experienced by the individual networks, specially for $\varepsilon$-greedy and EXP3. We identify the cause of this variability to be the adversarial setting of our setup in which the set of most played actions provide intermittent good/poor performance depending on the neighboring decisions. We also show that this variability is reduced using UCB and Thompson sampling, which are parameter-free policies that perform exploration according to the reward distribution of each action.
\end{abstract}

\begin{keyword}
High-Density Wireless Networks; Spatial Reuse; Resource Allocation; Decentralized learning; Multi Armed Bandits 
\end{keyword}

@@

%%%%%%%%%%%%%%%%%%%%%%%
%%%  I. INTRODUCTION 
%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{section:introduction}
Due to the growing popularity of wireless deployments, especially the ones based on the IEEE 802.11 standard (i.e., Wi-Fi), it is very common to find independent overlapping Wireless Networks (WNs) sharing the same channel resources. Due to the decentralized nature of such kind of deployments, the lack of organization and/or agreement on sharing policies leads to inefficient resources allocation, both in spectral and spatial domains. An illustrative example about the inefficiency noticed in WNs regarding channel sharing can be found in \cite{akella2007self}, where the authors show that the power level used by wireless devices is typically set, by default, to the maximum, regardless of the distance between communicating nodes, and of the channel occupancy \cite{akella2007self}. In decentralized environments, providing a model to obtain the optimal solution can be extremely complex. In WNs, the interactions between devices depend on many features (such as position, environment, transmit power) and are hard to derive. On top of this, network dynamics, in terms of user arrivals and departures, makes the resource allocation problem even harder.

In order to address the SR problem in decentralized WNs, we focus attention on Reinforcement Learning (RL), which has recently become very popular to solve many well-known problems in wireless communications. Some examples can be found for packet routing \cite{littman1993distributed}, Access Point (AP) selection \cite{bojovic2011supervised, bojovic2012neural}, optimal rate sampling \cite{combes2014optimal}, or energy harvesting in heterogeneous networks \cite{miozzo2015distributed}. Using online learning, a learner (or agent) obtains data in a sequential order and uses it to predict future good-performing actions. Online learning is, therefore, useful to cope with complex and dynamic environments. This background encourages us to approach a solution for the decentralized SR problem in WNs through online learning techniques. In particular, we exploit dynamic channel selection and transmit power adjustment for dealing with the performance issues resulting from uncoordinated resource sharing in WNs. While a proper frequency planning allows to reduce the interference between wireless devices, tuning the transmit power adds an extra level of SR that can result in improved throughput and fairness.

From the family of online algorithms, we propose the utilization of Multi-Armed Bandits (MABs) \cite{BCB12}, which is a well-known model in the online learning literature for solving resource allocation problems. In MABs, a given agent seeks to learn a hidden reward distribution while maximizing the gains. This is known as the exploration-exploitation trade-off. %(exploitation focuses on obtaining the maximum instantaneous reward, while exploration aims to maximize the long-term reward)	
While exploitation selects actions with the goal of maximizing long-term reward given the current estimate, exploration selects actions with the goal of improving the estimate. Unlike for classical RL, MABs do not consider states\footnote{A state refers to a particular situation experienced by a given agent, which is defined by a set of conditions. By having an accurate knowledge on its current situation, an agent can define state-specific strategies that maximize its profits.} in general, which can be hard to define for the decentralized SR problem. On the one hand, spatial interference cannot be binary treated, thus leading to complex interactions between nodes. On the other hand, the adversarial setting unleashed by decentralized deployments increases the system complexity, since the state not only depends on the actions taken by a given node, but also on the adversaries behavior.

This work extends our previous results presented in \cite{wilhelmi2017implications}, which were obtained by applying a stateless variation of Q-learning to the SR problem in WNs. Here we generalize the contributions done by implementing several action-selection strategies to allow WNs choose the channel and transmit power to use based on the resulting performance. In this paper, then, we provide a tutorial-based application of MABs to the decentralized SR problem, in order to maximize performance while observing the result from using different SR settings. According to that, we aim to illustrate the main benefits and drawbacks of applying different well-known learning techniques in a selfish way, i.e., independent WNs base their strategy on their own experienced performance. On the one hand, we evaluate the impact of varying parameters intrinsic to the proposed algorithms on the resulting throughput and fairness. In addition, we analyze the effects of learning selfishly, in order to shed some light on the future of decentralized approaches. Notably, we observe that even though players act selfishly, there are algorithms %learn to play collaborative actions
learn to play actions that enhance the overall performance, some times at the cost of high temporal variability. Considering selfish WNs and still obtaining collaborative behaviors is appealing to typical chaotic and dynamic deployments as we get rid of the need for message passing, or inference of neighboring conditions.

@@

%%%%%%%%%%%%%%%%%%%%%%%
%%%  II. RELATED WORK 
%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
\label{section:related_work} 		
In this paper we approach the decentralized SR problem through Dynamic Channel Allocation (DCA) and Transmit Power Control (TPC). While the former aims to allocate the available frequency range among potentially overlapping WNs, the latter attempts to increase the number of parallel transmissions by adjusting the power transmitted by each WN. While DCA operates at the frequency level, TPC deals with more complex spatial interactions. For that, an approach that dynamically tunes the transmit power may severely affect higher communication layers such as network (routing) or transport (congestion). TPC directly affects the transmission range, the data rate and the energy consumption, and it may create unidirectional links that unleash fairness issues. In fact, acknowledgments (ACKs) and RTS/CTS frames assume bidirectionality.

DCA has been extensively exploited from the centralized perspective, especially through techniques based on graph coloring \cite{riihijarvi2005frequency, mishra2005weighted}. Despite these kind of approaches allow to effectively reduce the interference between WNs, a certain degree of communication is required. Regarding decentralized methods, the authors in \cite{akl2007dynamic} propose a very simple approach in which each AP maintains an interference map of their neighbors, so that channel assignment is done through interference minimization. Unfortunately, the interactions between APs due to decentralization are not studied. Separately, \cite{chen2007improved} proposes two decentralized approaches that rely in the interference measured at both APs and STAs to calculate the best frequency channels for dynamic channel allocation. To do so, a WN, in addition to the interference sensed by its associated devices, considers other metrics such as the amount of traffic, so that some kind of coordination is required at the neighbor level (periodic reporting).

For the case of TPC, one of the first works in studying power management is \cite{elbatt2000power}, which shows that large improvements can be achieved if selecting the appropriate transmit power level in ad-hoc WNs. A centralized approach for TPC is presented in \cite{tang2014joint}, which performs power control and Rate Adaptation (RA) in subgroups of Wireless Local Area Networks (WLANs). The creation of clusters allows defining independent power levels between devices in the same group, which are useful to avoid asymmetric links. However, to represent all the possible combinations, graphs can become very large, specially in high-density deployments. Furthermore, we find a decentralized approach that relies on real-time channel measurements \cite{gandarillas2014dynamic}. The proposed mechanism (so called Dynamic Transmission Power Control) is based on a set of triggered thresholds that increase/decrease the transmit power according to the situation. The main problem is that thresholds are set empirically (based on simulations), so the mechanism may not always improve performance.

@@

To the best of our knowledge, there is very little related work on applying multi-armed bandit techniques to the problem of resource allocation in WNs. In \cite{coucheney2015multi}, the authors propose modeling a resource allocation problem in LTE networks through MABs. In particular, a set of Base Stations (BS) learn the best configuration of Resource Blocks (RBs) in a decentralized way. For that purpose, a variation of EXP3 (so-called Q-EXP3) is proposed, which allows to reduce the strategy set. Despite a regret bound is provided, it is subject to the fact that an optimal resource allocation exists, i.e., every BS obtains the necessary resources. In addition, a large number of iterations is required to find the optimal solution in a relatively small scenario, thus revealing the difficulties shown by decentralized settings.

More related to the problem proposed here, the authors in \cite{maghsudi2015joint} show a channel selection and power control approach in infrastructureless networks, which is modeled through bandits. In particular, two different strategies are provided to improve the performance of two Device to Device (D2D) users (each one composed by a transmitter and a receiver), which must learn the best channel and transmit power to be selected. Similarly to our problem, users do not have any knowledge on the channel or the other's configuration, so they rely on the experienced performance in order to find the best configuration. An extension of \cite{maghsudi2015joint} is provided by the same authors in \cite{maghsudi2015channel}, which includes a calibrated predictor (referred in the work as \textit{forecaster}) to infer the behavior of the other devices in order to counter act their actions. In each agent, the information of the forecaster is used to choose the highest-rewarding action with a certain probability, while the rest of actions are randomly selected. Henceforth, assuming that all the networks use a given strategy $\mathcal{X}$, fast convergence is ensured. Results show that channel resources are optimally distributed in a very short time frame through a fully decentralized algorithm that does not require any kind of coordination. Both aforementioned works rely in the existence of a unique Nash Equilibrium, which favors convergence. In contrast, in this article we aim to extend Bandits utilization to denser deployments, and, what is more important, to scenarios with limited available resources in which there is not a unique Nash Equilibrium (NE) that allows fast-convergence. Thus, we aim to capture the effects of applying selfish strategies in a decentralized way (i.e., agent $i$ follows a strategy $\mathcal{X}_i$ that does not consider the strategies of the others) and we also provide insight about the importance of past information for learning in dense WNs, which has not been studied before.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%III. MABs APPLICATION									    %	
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Multi-Armed Bandits for Improving Spatial Reuse in WNs}
\label{section:mabs}	
The decentralized resource allocation problem in WNs we aim to solve in this work can be modeled as an approximation of contextual bandits. Originally, the contextual bandits problem considers the existence of a context (or a state) that influences the action-selection process. As a consequence, the set of available strategies varies with the context and the probability distribution of a given reward depends both on the chosen actions and also on the context. In the resource allocation problem at hand, the context could include information such as the number of neighboring WNs, their position, their current channel selection, the transmission power they use as well as their action-selection strategies, among other. However, we are interested in the case where no context can be inferred given that such information is hard and costly to obtain in practice, especially in dynamic scenarios. Thus, we model the decentralized SR problem in WNs as an adversarial bandit problem, in which the actions of a given WN affect the pay-off distributions of the others' actions. Such model requires the existence of a NE with a pure strategy,\footnote{A pure strategy NE is conformed by a set of strategies and payoffs, so that no player can obtain further benefits by deviating from its strategy.} which does not always hold for unplanned deployments due to the scarcity of the available resources and the number of overlapping WNs. Understanding the implications derived from such an adversarial setting in the absence of a NE is one the main goals of this paper, which, to the best of our knowledge, has been barely considered in the previous literature.

@@

For simplicity, we consider that each WN is composed by an AP and a single STA, so that downlink traffic is held. Note that in typical uncoordinated wireless deployments (e.g., residential buildings), STAs are typically close to the AP to which they are associated. Thus, having several STAs associated to the same AP does not significantly impact the inter-WNs interference we study in this work. According to that, $\Gamma_{i}^*$ is computed as $B \cdot \log_{2}(1 + \text{SNR}_{i})$, where $B$ is the bandwidth, and $\text{SNR}_{i}$ is the Signal-to-Noise ratio experienced by the STA of $\text{WN}_i$. We consider this model because it allows to properly capture the interference generated between WNs, which is sought to be minimized.

@@

For the remainder of this work, we study the interactions between several WNs placed in a 3-D scenario that occur when applying MABs in a decentralized manner (with parameters described later in Section \ref{section:simulation_parameters}). For simplicity, we consider WNs to be composed by an Access Point (AP) transmitting to a single Station (STA) in a downlink manner. Additionally, to analyze the adversarial setting when WNs share scarce resources, we consider that the number of channels is equal to half the number of coexisting WNs, thus making the SR even more challenging.

@@

%%%%%%%%%%%%%%
% V. SIMULATIONS   %
%%%%%%%%%%%%%%
\section{Performance Evaluation}
\label{section:performance_evaluation}	
In this Section, we evaluate the performance of each action-selection strategy presented in Section \ref{section:mabs} when applied to the decentralized SR problem in WNs.\footnote{All of the source code used in this work open \cite{fwilhelmi2017code}, encouraging sharing of algorithms between contributors and providing the ability for people to improve on the work of others under the GNU General Public License v3.0} For that purpose, we first evaluate the $\varepsilon$-greedy, EXP3, UCB and Thompson sampling policies in a controlled environment. Without loss of generality, we consider a symmetric configuration in this scenario in order to analyze the competition effects when WNs have the same opportunities for accessing the channel. Then, in Section \ref{section:random}, we provide a performance comparison of the aforementioned scenarios with different densities and with randomly located WNs.

\subsection{Toy Grid Scenario}		
\label{section:toy_grid_scenario}	
We start considering a toy scenario consisting in a grid of 4 WNs, as illustrated in Figure \ref{fig:scenario}. This scenario has the particularity of being symmetric, so that adversarial WNs have the same opportunities to compete for the channel resources. Henceforth, this scenario allows us to \emph{i)} simplify the evaluation of the proposed algorithms, \emph{ii)} frame the competition between WNs on equal terms.

@@

Let actions range from $a_1$ to $a_8$, and map to \{channel number, transmit power (dBm)\}: $a_1$ = \{1, 5\}, $a_2$ = \{2, 5\}, $a_3$ = \{1, 10\}, $a_4$ = \{2, 10\}, $a_5$ = \{1, 15\}, $a_6$ = \{2, 15\}, $a_7$ = \{1, 20\} and $a_8$ = \{2, 20\}. According to our model, the joint configuration of the 4 WNs that grants the maximum aggregate throughput is any combination of actions \{$a_1$, $a_7$, $a_8$\}, subject to the condition of two close WNs using $a_1$. Despite the scenario is symmetric, the distance between APs is not the same in all the cases. For instance, $d_{\text{AP}_1,\text{AP}_2} > d_{\text{AP}_1,\text{AP}_4}$. It is important to emphasize that the proposed scenario does not consider asymmetries between WNs, so that dominance relationships are not present. In case of suffering from such asymmetries, the performance of a given WN would depend on the configuration of its neighboring networks but it won't be able to affect their behavior. For instance, a WN in the middle of two others may suffer from flow starvation if the WNs in the extremes do not sense each other. In such a situation, the WN in the middle is not able to fairly compete for channel resources due to the asymmetric situation it suffers.

In our symmetric scenario, the optimal aggregate throughput is 1124 Mbps, and is only achieved when two consecutive WNs sacrifice themselves by choosing a lower transmit power (action $a_1$). Moreover, to maximize the proportional fairness, the optimal joint configuration changes to any combination using the maximum transmit power (i.e., $a_7$ and $a_8$), so that frequency reuse is maximized; WNs using the same channel must be as much separated as possible. Such a configuration leads to an aggregate throughput of 891 Mbps, which is quite lower than in the former case, but allows experiencing a fairer configuration. Note that all WNs obtain the same throughput in this case as frequency reuse is achieved (closer WNs choose different channels), but power is kept equal for each WN and despite experiencing higher interference, none of the WNs sacrifice themselves. The aim of the following performance evaluation is to gain insight into whether a collaborative result, close to this equal division of resources given by the proportional fair action selection, is likely to be learned by the algorithms and if so, under which conditions.

>>>>

\begin{abstract}
Next-generation wireless deployments are characterized by being dense and uncoordinated, which often leads to inefficient use of resources and poor performance. To solve this, we envision the utilization of completely decentralized mechanisms to \francesc{enable} Spatial Reuse (SR). In particular, we concentrate in Reinforcement Learning (RL), and more specifically, in Multi-Armed Bandits (MABs), to allow networks to modify both their transmission power and channel based on their experienced throughput. In this work, we study the exploration-exploitation trade-off by means of the $\varepsilon$-greedy, EXP3, UCB and Thompson sampling action-selection strategies. Our results show that optimal proportional fairness can be achieved, even if no information about neighboring networks is available to the learners and WNs operate selfishly. However, there is high temporal variability in the throughput experienced by the individual networks, specially for $\varepsilon$-greedy and EXP3. We identify the cause of this variability to be the adversarial setting of our setup in which the set of most played actions provide intermittent good/poor performance depending on the neighboring decisions. We also show that this variability is reduced using UCB and Thompson sampling, which are parameter-free policies that perform exploration according to the reward distribution of each action.
\end{abstract}

\begin{keyword}
High-Density Wireless Networks, Interference, Decentralized learning, Multi Armed Bandits, Spatial Reuse, Resource Allocation.
\end{keyword}

@@

%%%%%%%%%%%%%%%%%%%%%%%
%%%  I. INTRODUCTION 
%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{section:introduction}
Due to the growing popularity of wireless deployments, especially the ones based on the IEEE 802.11 standard (i.e., Wi-Fi), it is very common to find independent overlapping Wireless Networks (WNs) \francesc{sharing the same channel resources. Due to the decentralized nature of such kind of deployments, the lack of organization and/or agreement on sharing policies leads to inefficient resources allocation, both in spectral and spatial domains. An illustrative example about the inefficiency noticed in WNs regarding channel sharing can be found in \cite{akella2007self}, where the authors show that the power level used by wireless devices is typically set, by default, to the maximum, regardless of the distance between communicating nodes, and of the channel occupancy \cite{akella2007self}. In decentralized environments}, providing a model to obtain the optimal solution can be extremely complex. In WNs, the interactions between devices depend on many features (such as position, environment, transmit power) and are hard to derive. On top of this, network dynamics, in terms of user arrivals and departures, makes the resource allocation problem even harder.

\francesc{In order to address the SR problem in decentralized WNs, we focus attention on Reinforcement Learning (RL), which has recently become very popular to solve many well-known problems in wireless communications. Some examples can be found for packet routing \cite{littman1993distributed}, Access Point (AP) selection \cite{bojovic2011supervised, bojovic2012neural}, optimal rate sampling \cite{combes2014optimal}, or energy harvesting in heterogeneous networks \cite{miozzo2015distributed}.} Using online learning, a learner (or agent) obtains data in a sequential order and uses it to predict future good-performing actions. Online learning is, therefore, useful to \francesc{cope with} complex and dynamic environments. This background encourages us to approach a solution for the decentralized SR problem in WNs through online learning techniques. \francesc{In particular, we exploit dynamic \francesc{channel selection} and transmit power adjustment \francesc{for dealing} with the performance issues resulting from uncoordinated resource sharing in WNs. While a proper frequency planning allows to reduce the interference between wireless devices, tuning the transmit power adds an extra level of SR that can result in improved throughput and fairness.}

From the family of online algorithms, we propose \francesc{the utilization of} Multi-Armed Bandits (MABs) \cite{BCB12}, which is a well-known model in the online learning literature for solving resource allocation problems. In MABs, \francesc{a given agent seeks} to learn a hidden reward distribution while maximizing the gains. This is known as the exploration-exploitation trade-off. %(exploitation focuses on obtaining the maximum instantaneous reward, while exploration aims to maximize the long-term reward)	
While exploitation selects actions with the goal of maximizing long-term reward given the current estimate, exploration selects actions with the goal of improving the estimate. \francesc{Unlike for classical RL, MABs do not consider states\footnote{\francesc{A state refers to a particular situation experienced by a given agent, which is defined by a set of conditions. By having an accurate knowledge on its current situation, an agent can define state-specific strategies that maximize its profits.}} in general, which can be hard to define for the decentralized SR problem. On the one hand, spatial interference cannot be binary treated, thus leading to complex interactions between nodes. On the other hand, the adversarial setting unleashed by decentralized deployments increases the system complexity, since the state not only depends on the actions taken by a given node, but also on the adversaries behavior.}

This work extends our previous results presented in \cite{wilhelmi2017implications}, which were obtained by applying a stateless variation of Q-learning to the SR problem in WNs. Here we generalize the contributions done by implementing several action-selection strategies to allow WNs choose the channel and transmit power to use based on the resulting performance. \francesc{In this paper, then, we provide a tutorial-based application of MABs to the decentralized SR problem, in order} to maximize performance while observing the result from using different SR settings. \francesc{According to that, we aim to illustrate the main benefits and drawbacks of applying different well-known learning techniques in a selfish way, i.e., independent WNs base their strategy on their own experienced performance. On the one hand, we evaluate the impact of varying parameters intrinsic to the proposed algorithms on the resulting throughput and fairness. In addition, we analyze the effects of learning selfishly, in order to shed some light on the future of decentralized approaches.} Notably, we observe that even though players act selfishly, there are algorithms %learn to play collaborative actions
learn to play actions that enhance the overall performance, some times at the cost of high temporal variability. Considering selfish WNs and still obtaining collaborative behaviors is appealing to typical chaotic and dynamic deployments as we get rid of the need for message passing, or inference of neighboring conditions. 
%\francesco{As later shown, this condition holds as long as adversarial WNs are able to compete for channel resources on equal terms. Otherwise, selfish learning would lead to unfairness situations, since WNs in a dominant position would take advantage of their situation, thus harming the performance of WNs in weaker situations.}
%	In particular, we exploit dynamic \francesc{channel selection} and transmit power adjustment \francesc{for dealing} with the performance issues resulting from uncoordinated resource sharing in WNs. While a proper frequency planning allows to reduce the interference between wireless devices, tuning the transmit power adds an extra level of SR that can result in improved throughput and fairness. %We aim to adjust both the channel allocation and transmit power by using local information only. This consideration models typical chaotic and dynamic deployments (e.g., residential buildings) without extra overhead (such as information sharing, additional software and/or hardware). 	
%	We consider selfish WNs that aim to maximize their own throughput and evaluate the impact of this on the resulting throughput and fairness while varying parameters intrinsic to the algorithms.

@@

%%%%%%%%%%%%%%%%%%%%%%%
%%%  II. RELATED WORK 
%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
\label{section:related_work} 		
\francesc{In this paper we approach the decentralized SR problem through Dynamic Channel Allocation (DCA) and Transmit Power Control (TPC). While the former aims to allocate the available frequency range among potentially overlapping WNs, the latter attempts to increase the number of parallel transmissions by adjusting the power transmitted by each WN. While DCA operates at the frequency level, TPC deals with more complex spatial interactions. For that, an approach that dynamically tunes the transmit power may severely affect higher communication layers such as network (routing) or transport (congestion). TPC directly affects the transmission range, the data rate and the energy consumption, and it may create unidirectional links that unleash fairness issues. In fact, acknowledgments (ACKs) and RTS/CTS frames assume bidirectionality.}

\francesc{DCA has been extensively exploited from the centralized perspective, especially through techniques based on graph coloring \cite{riihijarvi2005frequency, mishra2005weighted}. Despite these kind of approaches allow to effectively reduce the interference between WNs, a certain degree of communication is required. Regarding decentralized methods, the authors in \cite{akl2007dynamic} propose a very simple approach in which each AP maintains an interference map of their neighbors, so that channel assignment is done through interference minimization. Unfortunately, the interactions between APs due to decentralization are not studied. Separately, \cite{chen2007improved} proposes two decentralized approaches 
%and one centralized 
that rely in the interference measured at both APs and STAs to calculate the best frequency channels for dynamic channel allocation. To do so, a WN, in addition to the interference sensed by its associated devices, considers other metrics such as the amount of traffic, so that some kind of coordination is required at the neighbor level (periodic reporting).} %With respect to the centralized approach, which is shown to significantly improve the decentralized solutions, the main drawback is the requirement of a wired infrastructure for the management communication. Such deployment is often infeasible, especially in typical residential scenarios.

\francesc{For the case of TPC, one of the first works in studying power management is \cite{elbatt2000power}, which shows that large improvements can be achieved if selecting the appropriate transmit power level in ad-hoc WNs. A centralized approach for TPC is presented in \cite{tang2014joint}, which performs power control and Rate Adaptation (RA) in subgroups of Wireless Local Area Networks (WLANs). The creation of clusters allows defining independent power levels between devices in the same group, which are useful to avoid asymmetric links. However, to represent all the possible combinations, graphs can become very large, specially in high-density deployments. Furthermore, we find a decentralized approach that relies on real-time channel measurements \cite{gandarillas2014dynamic}. The proposed mechanism (so called Dynamic Transmission Power Control) is based on a set of triggered thresholds that increase/decrease the transmit power according to the situation. The main problem is that thresholds are set empirically (based on simulations), so the mechanism may not always improve performance.}

@@

To the best of our knowledge, there is very little related work on applying multi-armed bandit techniques to the problem of resource allocation in WNs. \francesc{In \cite{coucheney2015multi}, the authors propose modeling a resource allocation problem in LTE networks through MABs. In particular, a set of Base Stations (BS) learn the best configuration of Resource Blocks (RBs) in a decentralized way. For that purpose, a variation of EXP3 (so-called Q-EXP3) is proposed, which allows to reduce the strategy set. Despite a regret bound is provided, it is subject to the fact that an optimal resource allocation exists, i.e., every BS obtains the necessary resources. In addition, a large number of iterations is required to find the optimal solution in a relatively small scenario, thus revealing the difficulties shown by decentralized settings.}

\francesc{More related to the problem proposed here,} the authors in \cite{maghsudi2015joint} show a channel selection and power control approach in infrastructureless networks, which is modeled through bandits. In particular, two different strategies are provided to improve the performance of two Device to Device (D2D) users (each one composed by a transmitter and a receiver), which must learn the best channel and transmit power to be selected. Similarly to our problem, users do not have any knowledge on the channel or the other's configuration, so they rely on the experienced performance in order to find the best configuration. An extension of \cite{maghsudi2015joint} is provided by the same authors in \cite{maghsudi2015channel}, which includes a calibrated predictor (referred in the work as \textit{forecaster}) to infer the behavior of the other devices in order to counter act their actions. In each agent, the information of the forecaster is used to choose the highest-rewarding action with a certain probability, while the rest of actions are randomly selected. Henceforth, assuming that all the networks use a given strategy $\mathcal{X}$, fast convergence is ensured. Results show that channel resources are optimally distributed in a very short time frame through a fully decentralized algorithm that does not require any kind of coordination. Both aforementioned works rely in the existence of a unique Nash Equilibrium, which favors convergence. In contrast, in this article we aim to extend Bandits utilization to denser deployments, and, what is more important, to scenarios with limited available resources in which there is not a unique Nash Equilibrium (NE) that allows fast-convergence. Thus, we aim to capture the effects of applying selfish strategies in a decentralized way (i.e., agent $i$ follows a strategy $\mathcal{X}_i$ that does not consider the strategies of the others) and we also provide insight about the importance of past information for learning in dense WNs, which has not been studied before.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%III. MABs APPLICATION									    %	
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Multi-Armed Bandits for Improving Spatial Reuse in WNs}
\label{section:mabs}	
The decentralized resource allocation problem in WNs we aim to solve in this work can be modeled as an approximation of contextual bandits. Originally, the contextual bandits problem considers the existence of a context (or a state) that influences the action-selection process. As a consequence, the set of available strategies varies with the context and the probability distribution of a given reward depends both on the chosen actions and also on the context. In the resource allocation problem at hand, the context could include information such as the number of neighboring WNs, their position, their current channel selection, the transmission power they use as well as their action-selection strategies, among other. However, we are interested in the case where no context can be inferred given that such information is hard and costly to obtain in practice, especially in dynamic scenarios. Thus, we model the decentralized \francesc{SR} problem in WNs as an adversarial bandit problem, in which the actions of a given WN affect the pay-off distributions of the others' actions. Such model requires the existence of a NE with a pure strategy,\footnote{\francesc{A pure strategy NE is conformed by a set of strategies and payoffs, so that no player can obtain further benefits by deviating from its strategy.}} which \francesc{does not always hold for} unplanned deployments due to the scarcity of the available resources and the number of overlapping \francesc{WNs}. Understanding the implications derived from such an adversarial setting in the absence of a NE is one the main goals of this paper, \francesc{which, to the best of our knowledge, has been barely considered in the previous literature}.

@@

\francesc{For simplicity, we consider that each WN is composed by an AP and a single STA, so that downlink traffic is held. Note that in typical uncoordinated wireless deployments (e.g., residential buildings), STAs are typically close to the AP to which they are associated. Thus, having several STAs associated to the same AP does not significantly impact the inter-WNs interference we study in this work.} According to that, $\Gamma_{i}^*$ is computed as $B \cdot \log_{2}(1 + \text{SNR}_{i})$, where $B$ is the bandwidth, and $\text{SNR}_{i}$ is the Signal-to-Noise ratio experienced by the STA of $\text{WN}_i$. We consider this model because it allows to properly capture the interference generated between WNs, which is sought to be minimized.

@@

For the remainder of this work, \francesc{we study the interactions between several WNs placed in a 3-D scenario that occur when applying MABs in a decentralized manner} (with parameters described later in Section \ref{section:simulation_parameters}). For simplicity, we consider WNs to be composed by an Access Point (AP) transmitting to a single Station (STA) in a downlink manner. \francesc{Additionally, to analyze the adversarial setting when WNs share scarce resources,} we consider that the number of channels is equal to half the number of coexisting WNs, thus making the SR even more challenging.

@@

%%%%%%%%%%%%%%
% V. SIMULATIONS   %
%%%%%%%%%%%%%%
\section{Performance Evaluation}
\label{section:performance_evaluation}	
In this Section, we evaluate the performance of each action-selection strategy presented in Section \ref{section:mabs} when applied to the decentralized SR problem in WNs.\footnote{All of the source code used in this work open \cite{fwilhelmi2017code}, encouraging sharing of algorithms between contributors and providing the ability for people to improve on the work of others under the GNU General Public License v3.0} For that purpose, we first evaluate the $\varepsilon$-greedy, EXP3, UCB and Thompson sampling policies in a controlled environment. Without loss of generality, we consider a symmetric configuration in this scenario \francesc{in order to analyze the competition effects when WNs have the same opportunities for accessing the channel}. Then, in Section \ref{section:random}, we provide a performance comparison of the aforementioned scenarios with different densities and with randomly located WNs.

\subsection{Toy Grid Scenario}		
\label{section:toy_grid_scenario}	
\francesc{We start considering a toy scenario consisting in a grid of 4 WNs, as illustrated in Figure \ref{fig:scenario}. This scenario has the particularity of being symmetric, so that adversarial WNs have the same opportunities to compete for the channel resources. Henceforth, this scenario allows us to \emph{i)} simplify the evaluation of the proposed algorithms, \emph{ii)} frame the competition between WNs on equal terms.}

@@

Let actions range from $a_1$ to $a_8$, and map to \{channel number, transmit power (dBm)\}: $a_1$ = \{1, 5\}, $a_2$ = \{2, 5\}, $a_3$ = \{1, 10\}, $a_4$ = \{2, 10\}, $a_5$ = \{1, 15\}, $a_6$ = \{2, 15\}, $a_7$ = \{1, 20\} and $a_8$ = \{2, 20\}. \francesc{According to our model, the joint configuration of the 4 WNs that grants the maximum aggregate throughput is any combination of actions \{$a_1$, $a_7$, $a_8$\}, subject to the condition of two close WNs using $a_1$. Despite the scenario is symmetric, the distance between APs is not the same in all the cases. For instance, $d_{\text{AP}_1,\text{AP}_2} > d_{\text{AP}_1,\text{AP}_4}$.} \francesco{It is important to emphasize that the proposed scenario does not consider asymmetries between WNs, so that dominance relationships are not present. In case of suffering from such asymmetries, the performance of a given WN would depend on the configuration of its neighboring networks but it won't be able to affect their behavior. For instance, a WN in the middle of two others may suffer from flow starvation if the WNs in the extremes do not sense each other. In such a situation, the WN in the middle is not able to fairly compete for channel resources due to the asymmetric situation it suffers.}

\francesc{In our symmetric scenario, the optimal aggregate throughput is 1124 Mbps, and is only achieved when two consecutive WNs sacrifice themselves by choosing a lower transmit power (action $a_1$). Moreover, to maximize the proportional fairness, the optimal joint configuration changes to any combination using the maximum transmit power (i.e., $a_7$ and $a_8$), so that frequency reuse is maximized; WNs using the same channel must be as much separated as possible.} Such a configuration leads to an aggregate throughput of 891 Mbps, which is quite lower than in the former case, but allows experiencing a fairer configuration. Note that all WNs obtain the same throughput in this case as frequency reuse is achieved (closer WNs choose different channels), but power is kept equal for each WN and despite experiencing higher interference, none of the WNs sacrifice themselves. The aim of the following performance evaluation is to gain insight into whether a collaborative result, close to this equal division of resources given by the proportional fair action selection, is likely to be learned by the algorithms and if so, under which conditions.

%------------------------
% Study of MAB (e-greedy)
%------------------------
\subsubsection{Performance of the $\varepsilon$-greedy policy}
\label{section:egreedy_study}	
We first \francesc{evaluate} the performance of the $\varepsilon$-greedy policy when applied to the WNs problem, which allows us to study the purest form of the exploitation-exploration trade-off. First, we show the impact of modifying the initial exploration coefficient, $\varepsilon_0$ (recall that we have considered $\varepsilon_t = \varepsilon_0 / \sqrt{t}$), on the average throughput experienced by the overlapping WNs. Figure \ref{fig:egreedy_tuning_parameters} shows the aggregate throughput obtained in the grid scenario when applying $\varepsilon$-greedy during 10,000 iterations, and for each $\varepsilon_0$ value between 0 and 1 in 0.1 steps. The average and standard deviation of the throughput from 100 simulation runs are shown, which are compared with the optimal throughput presented in Section \ref{section:optimal_solution}.
\begin{figure}[t!]
	\centering							
	\epsfig{file=images/egreedy_tuning_parameters_zoom.eps, width=6.5cm}
	\caption{Average aggregate throughput and standard deviation obtained for each $\varepsilon_0$ value in $\varepsilon$-greedy. Results are from 100 simulations lasting 10,000 iterations each. The proportional fair solution is also shown (red dashed line).}
	\label{fig:egreedy_tuning_parameters}
\end{figure}		
As shown in the figure, the aggregate throughput obtained in average is quite similar for all the provided values of $\varepsilon_0$, except for the complete random case where no exploration is done ($\varepsilon_0 = 0$). The lower the $\varepsilon_0$ parameter, the less exploration is performed. Consequently, for low $\varepsilon_0$, the average throughput is highly dependent on how good/bad were the actions taken at the beginning of the learning process, which results in a higher standard deviation as $\varepsilon_0$ goes to 0.
